
---
title: "Cleaning bat acoustic data"
output: html_document
date: "2022-11-25"
---

##  Work environment set up 

Input: Results of the manual acoustic analysis for bat acoustic data collected at the Marker wind park. 
All acoustic analysis done by Reed April McKay. 
The dataset is a combination of all the csv files from each detector site location. 

Goal: 
Remove noise files, spelling errors, and simplify syntax for exploration and aggregation. 

Output: 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = "styler")


#### Load libraries ####
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
library(renv)
library(stringr)
library(janitor)

#### Import data  ####



manual_output <- read_csv("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs/Step1.cleaning bat acoustic data_2023-11-27/cleaed_manual.output_27.11.2023.csv") 
#257516 passes with noise, 34 vars 

summary(manual_output)

str(manual_output) #27516 obs of 47 variables - all recordings processed through Kaleidoscope pro combined from all sites, with file directories describing site origins and data collection times.


# for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"



## 
 file.name <- "Step1.cleaning bat acoustic data"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
 dir.create(dir.name)

output_today <- dir.name 

setwd(output_today)
getwd()


metacolumntext <- as.data.frame(names(manual_output)) 
#write.csv(metacolumntext, "column_names_1_Marker2020_ManualID_withnoise_original.csv")
# 16.11.2023 

```

# Remove noise, clean typos and simplify syntax
```{r}
manual_output <- mo
#### Remove noise files ####

manual_output$AUTO.ID. <- as.factor(manual_output$AUTO.ID.)

manual_output1 <- manual_output %>% filter(AUTO.ID. != "Noise")
# 19297 passes (not yet parsed for passes with multiple passes)
# 257516 - 19297 = 238219 noise files. 

#### Standardize manual ids ####


manual_output1$m_id <- as.factor(manual_output1$MANUAL.ID) 
summary(manual_output1$m_id)


# Edit a file name that should have been left in the raw format

fixthis1 <- subset(manual_output1, IN.FILE == "S4U10067_20200716_232445_000_SREwLRE.wav")
fixthis1$IN.FILE <- gsub("_SREwLRE", "", fixthis1$IN.FILE)
fixthis1$MANUAL.ID <- "SRE__LRE"

# drop the original observation from the dataset and replace it with the fixed observation 
manual_output2 <- manual_output1 %>% filter(IN.FILE != "S4U10067_20200716_232445_000_SREwLRE.wav") # dropped one observation 
manual_output3 <- full_join(manual_output2, fixthis1) # added back one observation 
# no more NAs in the MANUAL.ID column 

## This file S4U08483_20200808_002648.wav was tagged as a PAUR feeding buzz, but PAUR rarely show feeding buzzes. I revisited this file to check it and it is just a commuting PAUR that is perhaps approaching something, but I can't call that a feeding buzz. I need to change this to a commuting pass. 
fixthis2 <- subset(manual_output1, IN.FILE == "S4U08483_20200808_002648_000.wav")
fixthis2$MANUAL.ID  <- gsub(".B", "", fixthis2$MANUAL.ID)

manual_output3x <- manual_output3 %>% filter(IN.FILE != "S4U08483_20200808_002648_000.wav")

manual_output3xx <- full_join(manual_output3x, fixthis2)

# The barbastelle recordings (BABA) were sent to external validators to assess species ID
# 10 out of 11 were validated, one was found to be a poor recording of either a Myotis species or potential Plecotus auritus and will now be designated as (SR1)
# file name: MARKER-2020_20200901_230046_000.wav 

fixthis3 <- subset(manual_output3, IN.FILE == "MARKER-2020_20200901_230046_000.wav")
fixthis3$MANUAL.ID <- gsub("BABA", "SRE", fixthis3$MANUAL.ID)
fixthis3$m_id <- gsub("BABA", "SRE", fixthis3$m_id)
fixthis3

# drop the original observation from the dataset and replace it with the fixed observation 
manual_output3x <- manual_output3 %>% filter(IN.FILE != "MARKER-2020_20200901_230046_000.wav") # dropped one observation 
manual_output3xx <- full_join(manual_output3x, fixthis3) 
   
## Had to make these corrections unfortunately AFTER I had already prepared the whole script, so I am going to reuse the name manual_output3 to keep this simple... 
manual_output3 <- manual_output3xx


## More cleaning 

# change __ to :
manual_output3$m_id <- as.factor(manual_output3$MANUAL.ID) 
summary(manual_output3$m_id)
manual_output3$m_id1<- gsub("__", ":", manual_output3$m_id )
# the guild preface is redundant and can be removed. 
manual_output3$m_id2 <- gsub("LRE_","", manual_output3$m_id1)
manual_output3$m_id3 <- gsub("MRE_","", manual_output3$m_id2)
manual_output3$m_id4 <- gsub("SRE_","", manual_output3$m_id3)
# weird typo
manual_output3$m_id5 <- gsub("EMU","VEMU", manual_output3$m_id4) # this will create some more weird typos
manual_output3$m_id6 <- gsub("VVEMU","VEMU", manual_output3$m_id5)
# make one all encompassing NoID column 
manual_output3$m_id7 <- gsub("UNK_BABA_EPNI","NoID", manual_output3$m_id6) 
manual_output3$m_id8 <- gsub("UNK","NoID", manual_output3$m_id7) 
manual_output3$m_id9 <- gsub("_UNK","NoID", manual_output3$m_id8)

# some more typos
manual_output3$m_id10 <- gsub("LRE1","LR1", manual_output3$m_id9)
# simplify the social call tags
manual_output3$m_id11 <- gsub(".A/D",".S", manual_output3$m_id10)
manual_output3$m_id12 <- gsub(".S/D",".S", manual_output3$m_id11)
manual_output3$m_id13 <- gsub(".S/A",".S", manual_output3$m_id12)

# Anything tagged as PIPI or potentially PI will become MR1
# MR2 also becomes MR1 (unknown Pipistrelle calls)
manual_output3$m_id14 <- gsub("PIPI","MR1", manual_output3$m_id13)
manual_output3$m_id15 <- gsub("MR2","MR1", manual_output3$m_id14)
manual_output3$m_id16 <- gsub("MRE1","MR1", manual_output3$m_id15)
manual_output3$m_id17 <- gsub("MRE2","MR1", manual_output3$m_id16)
# lets check and see what this looks like now... 
manual_output3$m_id17 <- as.factor(manual_output3$m_id17)
summary(manual_output3$m_id17)
levels((manual_output3$m_id17))

# just a few more weird things to fix... 
manual_output3$m_id18 <- gsub("_NoID.S","NoID.S", manual_output3$m_id17)
manual_output3$m_id18 <- as.factor(manual_output3$m_id18)
summary(manual_output3$m_id18)
levels((manual_output3$m_id18)) # much better

### Now parse by multiple bats in a single passes 
mcols <- c("manual_pass1", "manual_pass2")

# Trim down to just essential columns
manual_output4 <- manual_output3 %>% 
  select(INDIR, OUTDIR, IN.FILE, DURATION, 
         DATE, TIME, HOUR, 
         DATE.12, TIME.12, HOUR.12, 
         AUTO.ID., PULSES, MATCH.RATIO, 
         MANUAL.ID, m_id18) #1927 obs of 15 vars

summary(manual_output4)

#write.csv(manual_output4, file.path(output_today, "Marker_manualIDscleanred_preparsed.csv")) 

manual_output5 <- manual_output4 %>% 
  tidyr::separate( col = m_id18,
                   sep = ":", 
                   into = mcols)

# create a subset of the dataset with just manual_pass2, 
# remove the NAs
# add back these observations as new passes

bat2 <- manual_output5 %>% select(INDIR, OUTDIR, IN.FILE, DURATION, 
         DATE, TIME, HOUR, 
         DATE.12, TIME.12, HOUR.12, 
         AUTO.ID., PULSES, MATCH.RATIO, 
         MANUAL.ID, manual_pass2) %>% 
  mutate(manual_pass2 = factor(manual_pass2)) %>% 
  rename(manual_pass_clean1 = manual_pass2) # need to have all the same column names to merge
summary(bat2) # 19111 NAs in manual_pass2

bat2.0 <- drop_na(bat2) # 186 obs 
#19297 - 186 = 19111 # good! 

```

## Parsing passes for multiple bats in one pass and creating the behavior column from manual.id codes
```{r}

#### Parsing multiple bats from passes ####

# make a dataset with just manual pass1 
bat1 <- manual_output5 %>% select(INDIR, OUTDIR, IN.FILE, DURATION, 
         DATE, TIME, HOUR, 
         DATE.12, TIME.12, HOUR.12, 
         AUTO.ID., PULSES, MATCH.RATIO, 
         MANUAL.ID, manual_pass1) %>% 
  mutate(manual_pass1 = factor(manual_pass1)) %>% 
  rename(manual_pass_clean1 = manual_pass1) # need to have all the same column names to merge

manual_output6 <-full_join(bat1, bat2.0) #19483 obs 
#19297 + 186 = 19483 # good! 

summary(manual_output6)


#### Parse out behavior ####

# make behavior column, remove behavior tags from manuals ids 
manual_output7 <- manual_output6 %>%
  mutate(manual_pass_clean1 = as.character(manual_pass_clean1)) %>% 
  mutate(behavior = case_when(
    str_detect(manual_pass_clean1, ".B") ~ "Feeding",
    str_detect(manual_pass_clean1, ".S") ~ "Social",
    str_detect(manual_pass_clean1,".B.S") ~ "Feeding",
    str_detect(manual_pass_clean1, ".S.B") ~ "Feeding",
    TRUE ~ "Commuting"
  )) 

# manual_output7$manual_pass_clean1 <- as.factor(manual_output7$manual_pass_clean1)
# summary(manual_output7$manual_pass_clean1)
# All of the BABA passes were defined as Feeding when this happened - all of these passes were just commuting so I need to fix that. 

manual_output8 <-manual_output7 %>% 
  mutate(behavior = case_when(
  MANUAL.ID == "SRE_BABA" ~ "Commuting", 
  TRUE ~ as.character(behavior))) 

## 
manual_output8$behavior <- as.factor(manual_output8$behavior) 
   summary(manual_output8$behavior)
  # Commuting   Feeding    Social 
  #  14887      4046       550 
   
   
  
# remove all the behavior tags in the manual_pass_clean1 column so it is only manual ids
manual_output8$manual_pass_clean2 <- gsub(".B", "",manual_output8$manual_pass_clean1)
manual_output8$manual_pass_clean3 <- gsub(".S", "",manual_output8$manual_pass_clean2)

manual_output8$manual_pass_clean3 <- as.factor(manual_output8$manual_pass_clean3)
summary(manual_output8$manual_pass_clean3)
#   BA  EPNI EPNI.   LR1   LR2   LRE   MR1  NYNO  NoID Noise  PAUR  PINA  PIPY   SRE  SRE.  VEMU VEMU. 
#   10 14014     1   131   221     1    22   182   484    45   553     9    64  3577     1   167     1 

# Okay SO still some typos in here but nothing catastrophic!
manual_output8$manual_pass_clean4 <- gsub("BA", "BABA", manual_output8$manual_pass_clean3)
manual_output8$manual_pass_clean5 <- gsub("EPNI.", "EPNI", manual_output8$manual_pass_clean4)
manual_output8$manual_pass_clean6 <- gsub("LRE", "LR1", manual_output8$manual_pass_clean5)
manual_output8$manual_pass_clean7 <- gsub("SRE.", "SRE", manual_output8$manual_pass_clean6)
manual_output8$manual_pass_clean8 <- gsub("SRE", "SR1", manual_output8$manual_pass_clean7) 
manual_output8$manual_pass_clean9 <- gsub("VEMU.", "VEMU", manual_output8$manual_pass_clean8)

manual_output8$manual_pass_clean9 <- as.factor(manual_output8$manual_pass_clean9)
summary(manual_output8$manual_pass_clean9) # WOOOOHOOOOO
 # BABA  EPNI   LR1   LR2   MR1  NYNO  NoID Noise  PAUR  PINA  PIPY   SR1  VEMU 
 #   11 14015   132   221    22   182   484    45   553     9    64  3578   168

manual_output8$behavior <- as.factor(manual_output8$behavior)
summary(manual_output8$behavior)
# Commuting   Feeding    Social 
#     14887     4046       550 


manual_output9 <- manual_output8 %>% select(INDIR, OUTDIR, IN.FILE, DURATION, 
         DATE, TIME, HOUR, 
         DATE.12, TIME.12, HOUR.12, 
         AUTO.ID., PULSES, MATCH.RATIO, 
         MANUAL.ID, manual_pass_clean9, behavior) %>% 
         rename(manual.id = manual_pass_clean9) 

# Export before removing noise 
#write.csv(manual_output9, "MarkerManualIDs_cleaned_withNoise.csv")


#### Drop Noise files  ####

manual_output10 <- manual_output9 %>% filter(manual.id != "Noise") #19438 passes
#19483-19438 = 45 # good! 


#### Make guild column  ####

 # BABA  EPNI   LR1   LR2   MR1  NYNO  NoID PAUR  PINA  PIPY   SR1  VEMU 
manual_output11 <- manual_output10 %>% mutate(
  guild = case_when(
    manual.id %in% c("BABA", "SR1", "PAUR") ~ "SRE", 
    manual.id %in% c("MR1", "PINA", "PIPY") ~ "MRE", 
    manual.id %in% c("EPNI", "LR1", "LR2", "VEMU", "NYNO") ~ "LRE",
    manual.id == "NoID" ~ "NoID"
    
  )
) %>% mutate(guild = as.factor(guild))
 
summary(manual_output11$guild)
#   LRE   MRE  NoID   SRE 
# 14718    95   484  4141 


#### Make the Site / Habitat / Locality / Facility columns ####

#summary(compare)
head(manual_output11)

manual_output12 <- manual_output11 %>%
  mutate(Site = case_when(
    str_detect(OUTDIR, "MetA") ~ "Met45",
    str_detect(OUTDIR, "MetB") ~ "Met95",
    str_detect(OUTDIR, "C02") ~ "N02",
    str_detect(OUTDIR, "P02") ~ "P02",
    str_detect(OUTDIR, "C04") ~ "N04",
    str_detect(OUTDIR, "P04") ~ "P04",
    str_detect(OUTDIR, "C08") ~ "N08",
    str_detect(OUTDIR, "P08") ~ "P08",
    str_detect(OUTDIR, "C09") ~ "N09",
    str_detect(OUTDIR, "P09") ~ "P09",
    str_detect(OUTDIR, "C10") ~ "N10",
    str_detect(OUTDIR, "P10") ~ "P10",
    str_detect(OUTDIR, "C11") ~ "N11",
    str_detect(OUTDIR, "P11") ~ "P11",
    str_detect(OUTDIR, "C14") ~ "N14",
    str_detect(OUTDIR, "P14") ~ "P14",
  )) %>% mutate(Site = as.factor(Site)) 

summary(manual_output12$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#   197    35   790   600   598  4287  1819   368   728   836   783   441  3686   976  1835  1459 

manual_output13 <- manual_output12 %>% mutate(Habitat = 
                   case_when(
                     str_detect(Site, "Met") ~ "MeteorologicalTower", 
                     str_detect(Site, "N") ~ "Natural", 
                     str_detect(Site, "P") ~ "TurbinePad", 
                   )) %>%  mutate(Habitat = as.factor(Habitat)) 

summary(manual_output13$Habitat)
# MeteorologicalTower             Natural          TurbinePad 
#                 232                9190               10016 
summary(manual_output13$Site)
manual_output14 <- manual_output13 %>% mutate(Locality = 
                   case_when(
                     str_detect(Site, "Met") ~ "MeteorologicalTower", 
                     str_detect(Site, "2") ~ "Turbine2", 
                     Site %in% c("P04", "N04") ~ "Turbine4", 
                     str_detect(Site, "8") ~ "Turbine8", 
                     str_detect(Site, "9") ~ "Turbine9", 
                     str_detect(Site, "10") ~ "Turbine10", 
                     str_detect(Site, "11") ~ "Turbine11", 
                     Site %in% c("P14", "N14") ~ "Turbine14",
                   ))

manual_output14 <- manual_output14 %>%  mutate(Locality = as.factor(Locality)) 
summary(manual_output14$Locality)
# MeteorologicalTower           Turbine10           Turbine11           Turbine14            Turbine2 
#                 232                2795                2203                2187                1626 
#            Turbine4            Turbine8            Turbine9 
#                1383                1039                7973 

manual_output15 <- manual_output14 %>% mutate(Facility = 
                   case_when(
                     str_detect(Site, "Met") ~ "South", 
                     str_detect(Site, "2") ~ "North", 
                     str_detect(Site, "4") ~ "North", 
                     str_detect(Site, "8") ~ "North", 
                     str_detect(Site, "9") ~ "South", 
                     str_detect(Site, "10") ~ "South", 
                     str_detect(Site, "11") ~ "South", 
                     str_detect(Site, "14") ~ "South", 
                   )) %>%  mutate(Facility = as.factor(Facility)) 
summary(manual_output15$Facility)
# North South 
# 6235 13203 

test <- manual_output15 %>% 
  group_by(guild, behavior) %>% 
  dplyr::summarise(n = n())
sum(test$n)

kable(test) %>% kable_minimal()

summary(manual_output15)
getwd()

```


## Exploratory plots to check that the data looks alright 
```{r}

names(manual_output15)
#  [1] "INDIR"       "OUTDIR"      "IN.FILE"     "DURATION"    "DATE"        "TIME"        "HOUR"        "DATE.12"     "TIME.12"     "HOUR.12"    
# [11] "AUTO.ID."    "PULSES"      "MATCH.RATIO" "MANUAL.ID"   "manual.id"   "behavior"    "guild"       "Site"        "Habitat"     "Locality"   
# [21] "Facility"  

# are all the guilds represented at each site throughout the season? 
ggplot(manual_output15) + geom_point(aes(x=DATE, y = guild, color = guild)) + facet_wrap(~Site) # yes
# are all the guilds represented at each site throughout the night? 
ggplot(manual_output15) + geom_point(aes(x=HOUR.12, y = guild, color = guild)) + facet_wrap(~Site) # yes 

ggplot(manual_output15) + geom_count(aes(x=DATE, y = guild, color = guild)) + facet_wrap(~Site) # first plot but with geom_count instead of geom_point


### Export final output 

#write.csv(manual_output15, file.path(output_today,"MarkerManual_cleaned_manualid_guild_sitenames.csv"))
getwd()
```






