
---
title: "3. Aggregate bat acoustic data to hour with weather and insects"
output: html_document
date: "2023-01-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Input :  The cleaned manual acoustic analysis data from the Marker 2020 study 

Outputs: 1. All bat activity aggregated to hour per site - for total bat activity 
           
# Work environment set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(tidy = "styler")
options(knitr.table.format = "html")


library(knitr)
library(data.table)
library(tidyverse)
library(beepr)
library(lubridate)
library(purrr)
library(renv)
library(stringr)
library(janitor)
library(anytime)
library(kableExtra)
library(papeR)
library(skimr)
library(vtable)
library(suncalc)


#### Import data  ####

input <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Input"

dataset1 <- "MarkerManual_cleaned_manualid_guild_sitenames.csv"
dataset2 <- "Marker_night_inventory_0704.2022_edit.csv"
dataset3 <- "raw insect data tidied (1).csv"
dataset4 <- "TRUE_hourlyaverageweather_wholepark.csv"
# 
path1 <- str_c(input, "/", dataset1)
path2 <- str_c(input, "/", dataset2)
path3 <- str_c(input, "/", dataset3)
path4 <- str_c(input, "/", dataset4)
# 
bats1 <- read_csv(path1) # 19438 bat passes with 22 vars
active.nights <- read.csv(path2, sep = ";") # # 951 obs of 3 variables
bugs <- read.csv(path3) # 11420 obs 8 vars
weather.hours <- read_csv(path4) # 93 obs of 8 vars
# 
# 
# # for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"

## 
 file.name <- "3. Aggregate bat acoustic data to hour with weather and insects"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
 dir.create(dir.name) # be careful not to run over an already created directory 

output_today <- dir.name

setwd(output_today)

getwd()

```


# Summary tables with the unaggregated dataset 
```{r}
# make factor levels arranged in orders that are more meaningful and insert a taxa column with more specific manual id names. 
names(bats1)
cols <- c("manual.id" ,  "behavior"  ,  "guild"     ,  
          "Site"     ,   "Habitat"   ,  "Locality"  ,  "Facility")

bats1[cols] <- lapply(bats1[cols], factor)
bats1$taxa <- bats1$manual.id
levels(bats1$taxa)
# "BABA" "EPNI" "LR1"  "LR2"  "MR1"  "NYNO" "NoID" "PAUR" "PINA" "PIPY" "SR1"  "VEMU"
levels(bats1$taxa) <- list("Barbastella barbastellus" = "BABA",
                           "Plecotus auritus" = "PAUR",
                           "Short range group 1" = "SR1",
                           "Eptesicus nilssonii" = "EPNI",
                           "Nyctalus noctula" = "NYNO",
                           "Vespertilio murinus" = "VEMU",
                           "Long range group 1" = "LR1",
                           "Long range group 2" = "LR2",
                           "Medium range group1" = "MR1",
                           "Pipistrellus nathusii" = "PINA",
                           "Pipistrellus pygmaeus" = "PIPY",
                            "Unknown bat" = "NoID") 
summary(bats1$taxa)

levels(bats1$guild) <- list("SRE" = "SRE", "MRE" = "MRE", "LRE" = "LRE", "NoID" = "NoID")
summary(bats1$guild)

levels(bats1$Locality)
levels(bats1$Locality) <-list("Turbine2" = "Turbine2" , "Turbine4" = "Turbine4",  "Turbine8" = "Turbine8", "MeteorologicalTower" = "MeteorologicalTower", "Turbine9" = "Turbine9", "Turbine10" = "Turbine10", "Turbine11" = "Turbine11", "Turbine14" = "Turbine14") 
summary(bats1$Locality)

# Can play with these and make them much nicer... 
kbl(summarize(bats1, type = "factor", variables = "manual.id"))  %>%  kable_styling()
kbl(summarize(bats1, type = "factor", variables = "taxa"))  %>%  kable_styling()
kbl(summarize(bats1, type = "factor", variables = "guild"))  %>%  kable_styling()
kbl(summarize(bats1, type = "factor", variables = "Site"))  %>%  kable_styling()
kbl(summarize(bats1, type = "factor", variables = "Habitat"))  %>%  kable_styling()
kbl(summarize(bats1, type = "factor", variables = "Facility"))  %>%  kable_styling()

names(bats1)
bats_st <- bats1 %>% select(HOUR.12, HOUR, DATE, DATE.12, behavior, guild, Site, Habitat, Locality, Facility, taxa) 

st(bats_st) # summary table of all the relevant factor levels across the whole dataset

# export the html file: 

# st(bats2., file = file.path(output_today, "total summary table"))

```


# Aggregate total ground level bat acoustic activity by hour 
- Insert zeros
- merge with night aggregated weather data 
```{r}

# create night-hour column
bats1$nighthour <- paste(bats1$DATE.12, " ",bats1$HOUR.12, ":00", sep= "")
bats1$datehour <- paste(bats1$DATE, " ",bats1$HOUR, ":00", sep= "")

bats2 <- bats1 %>% mutate(date = as.Date(DATE),
                           night = as.Date(DATE.12),
                           nighthour = ymd_hm(nighthour),
                           datehour = ymd_hm(datehour))  

#### CHECK THE TIME ZONE ####  
tz(bats2$nighthour)
tz(bats2$datehour)

bats2$datehour <- force_tz(bats2$datehour, tzone = "Europe/Berlin")
tz(bats2$datehour)

bats2$nighthour <- force_tz(bats2$nighthour, tzone = "Europe/Berlin")
tz(bats2$nighthour)


# Locations map, to merge back later 
bat_sitemap <- bats1 %>% select(Site, Habitat, Facility, Locality) %>% distinct() 

# Aggregate by datehour 
bats_hours_sum <- bats2 %>% group_by(Site, datehour) %>% dplyr::summarize(batpass = sum(n())) 
bats_hours_sum1 <- left_join(bats_hours_sum, bat_sitemap)

summary(bats_hours_sum1)
summary(bats_hours_sum1$Site)

dat <- bats_hours_sum1 %>% mutate(date = as.Date(datehour)) 
summary(dat)
summary(dat$Site)

# Time map, to merge back later 
timemap <- bats1 %>% select(Site, datehour, nighthour, DATE, DATE.12, HOUR, HOUR.12) %>% distinct() 

timemap$datehour <- ymd_hm(timemap$datehour)
timemap$nighthour <- ymd_hm(timemap$nighthour)


timemap$datehour <- force_tz(timemap$datehour, tzone = "Europe/Berlin")
tz(timemap$datehour)

timemap$nighthour <- force_tz(timemap$nighthour, tzone = "Europe/Berlin")
tz(timemap$nighthour)

summary(timemap$Site)
## The number of hours with recorded bat passes per site 
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#  127    24   221   157   199   232   259   192   235   203   187   166   257   245   259   219 

#write.csv(bats_nights_sum1, "bat pass aggregated by night wout zeronights.csv")

################################################
# Insert zero activity nights
################################################
head(active.nights)

active.nights$Site <- gsub("C", "N", active.nights$Site)
active.nights$Site <- gsub("MetA", "Met45", active.nights$Site)
active.nights$Site <- gsub("MetB", "Met95", active.nights$Site)
active.nights$Site <- as.factor(active.nights$Site)
active.nights$night <- as.Date(active.nights$night, '%d.%m.%Y')
active.nights$night <- as.Date(active.nights$night)
summary(active.nights)

#---- ADDING MISSING DAYS AND HOURS

# adjust dataset to include nights when detectors were active but no data was collected. 

head(active.nights)
summary(active.nights$Site)

nights <- unique(active.nights$night) # 91 unique nights of survey

# restrict to only between one hour before sunset and one hour after sunrise 
## Recalculate day_length 

# manually add 09.30 
thenights <- as.data.frame(active.nights$night %>% unique())
colnames(thenights)[1] = "date"
day0930 <- as.data.frame(as.Date("2020-09-30")) 
colnames(day0930 )[1] = "date"
thenights1 <- full_join(thenights, day0930)
# Will have to use night as a proxy for date


#Use location of T-08
photoperiod <- getSunlightTimes(
  date = thenights1$date, 
  keep = c("sunrise", "sunset"),
  lat = 59.49964,
  lon = 11.73199, 
  tz = "Europe/Berlin") # MAKE SURE THIS MATCHES YOUR OTHER DATA 

photoperiod1 <- photoperiod %>%
  mutate(
    date = as.POSIXct(date),
    day_length = as.numeric(sunset - sunrise)) %>% distinct()

photoperiod2 <- photoperiod1 %>% mutate(night = as.Date(date)) %>% select(-date) %>% distinct()
summary(photoperiod2)
# write.csv(photoperiod2, "day_length for active.nights Marker2020.csv")


# Calculate the active hours for each site 
# Make 24 hours for each active.night
hours.expand <- expand.grid(night = active.nights$night, hour = sprintf("%02d:00", 0:23)) 
active.nights1 <- active.nights %>% select(c(Site, night)) 
hours.expand1 <- left_join(active.nights, hours.expand, by = "night") %>% distinct() 
hours.expand1$nighthour <- paste(hours.expand1$night, hours.expand1$hour)
hours.expand1$nighthour <- ymd_hm(hours.expand1$nighthour)
hours.expand2 <- hours.expand1 %>% select(Site, nighthour) %>% distinct()
str(hours.expand2)

# 22824 obs = 951 detector nights * 24 hours - good! 

nighthours <- hours.expand2
summary(nighthours)
#write.csv(hours.expand2, "24active.nighthours.Marker2020.csv")

# merge the bats dataset to with the photoperiod dataset 
photoperiod1$date <- as.Date(photoperiod1$date)
dat1 <- left_join(dat, timemap) %>% select(-c(Habitat, Locality, Facility, datehour, date, DATE, DATE.12, HOUR, HOUR.12)) %>% distinct()
summary(dat1)
str(dat1)
str(hours.expand2)

dat2 <- merge(dat1, hours.expand2, all.y = TRUE) %>% distinct()
summary(dat2) # 19642  NAs introduced 
#22824 obs 

dat3 <- dat2 %>% mutate_at(3, ~replace_na(.,0))  
sum(dat3$batpass) # 19438 - good! 

# add back date and datehour then photoperiods/sunset data
dat4 <- dat3 %>% mutate(datehour = nighthour + 12*60*60,
                        date = as.Date(datehour))

dat5 <- left_join(dat4, photoperiod1) %>% distinct()

dat6 <- dat5 %>% mutate(starttime = sunset - hours(2), endtime = sunrise + hours(1)) 
summary(dat6) 

dat7<- dat6%>%  
  mutate(Keep = case_when(
  datehour >= starttime | datehour <= endtime ~ "Yes", TRUE ~ "No"
))

dat7$Keep <- as.factor(dat7$Keep)
summary(dat7) # 8981 observations that do not fit. 

#Do any have bat passes? 
test <- dat7 %>% filter(Keep == "No")
sum(test$batpass) 

# would lose 15 bat passes if I kept the cut off at 1 hour before sunset 
# (no issues with the 1 hour cutoff at sunrise)... 
# Adjusting the cut off to 2 hours before sunset gives us more zeros in the dataset but keeps all bat passes.

dat8<- dat7 %>% filter(Keep == "Yes") %>% select(-c(Keep)) %>% droplevels() 
summary(dat8)

# 13843 detector hours for total bat passes 
# NOW: 10990 detector hours 

# write.csv(dat8, "hourlyaggregatedMarker2020_totalbats_zeroinserted.csv") 

## Add back location information and julian night 

dat9 <- left_join(dat8, bat_sitemap)
summary(dat9)
dat9$night <- as.Date(dat9$nighthour)
dat9$jnight <- yday(dat9$night)

```


# Combine with weather data by night hour 
```{r}
head(weather.hours)

dat10 <- left_join(dat9, weather.hours, by = "nighthour") %>% select(- ...1)

summary(dat10)


sum(dat10$batpass) # 19438

# 32 NA observations added - do they include bat passes? 

datx <- dat10 %>% 
  select(-c(night.y,  date.y, datehour.y, jnight.y)) %>% 
  mutate(night = night.x, 
         date = date.x, 
         datehour = datehour.x,
         jnight = jnight.x) %>% 
  select(-c(night.x, date.x, datehour.x, jnight.x))

datx1 <- na.omit(datx)
sum(datx1$batpass)
19438 - 19396 # lose 42 bat passes in the process... 

summary(datx1) # 10958 observations
sum(datx1$batpass) # 19396
getwd()
#write.csv(datx1, "hourlyaggregated_totalbats_zeroinserted_weather.csv")
```



######################################

# *For aggregating with Insect data* 

######################################



# Aggregate bat data by behavior and site 
```{r}

# Locations map, to merge back later 
str(bat_sitemap)


# Some slight adjustments needed for merging later: 

active.hours <- dat5 %>% select(Site, nighthour) #
active.hours$active.hours <- "TRUE"
summary(active.hours$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#  1680  1680  1416  1056  1440  1152  1440  1440  1440  1440  1440  1440  1440  1440  1440  1440 

# Make a behavior row for each detector hour - for merging with the zero inserted dataset later. 
F.active.hours <- active.hours %>% mutate(behavior = "Feeding") 
S.active.hours <- active.hours %>% mutate(behavior = "Social") 
C.active.hours <- active.hours %>% mutate(behavior = "Commuting") 
active.hours1 <- full_join(F.active.hours, S.active.hours)
active.hours2 <- full_join(active.hours1, C.active.hours) # 68472 observations 
summary(active.hours2)

active.hours2$behavior <- as.factor(active.hours2$behavior)
summary(active.hours2$behavior)
# Commuting   Feeding    Social 
#    22824     22824     22824 

summary(active.hours2$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#  5040  5040  4248  3168  4320  3456  4320  4320  4320  4320  4320  4320  4320  4320  4320  4320 

#write.csv(active.hours2, "active.nighthours_trimmed-2SS+SR_eachbehavior.csv")

active.hours.b <- active.hours2
tz(active.hours$nighthour) # "Europe/Berlin" 

# rename dataset and check time zones 
dfb <- bats2
tz(dfb$nighthour)
tz(dfb$datehour)

# Aggregate 
dfb1 <- dfb %>% group_by(Site, behavior, nighthour) %>% dplyr::summarize(batpass = sum(n())) 

summary(dfb1)
summary(dfb1$behavior)
# Commuting   Feeding    Social 
     # 2989       718       368

sum(dfb1$batpass) #19438


#---- ADDING MISSING HOURS
# remember dat5 includes all active detector nights per Site
newnights <- active.hours.b %>% select(Site, nighthour, behavior) %>% 
  mutate(night = as.Date(nighthour),
         active.night = "TRUE") %>% select(-nighthour) %>% distinct()
head(newnights) #951 nights 

# Use existing dataset hours to create nighthour column for now. 
hours<-seq(min(dfb1$nighthour),max(dfb1$nighthour),3600) 
#--full daily sequence of hours for the entire study period

hours<-hours[as.numeric(format(hours,"%H"))%in%sort(unique(as.numeric(format(dfb1$nighthour,"%H"))))]
#--restrict to "nighthours" found in the data

temp<-expand.grid(Site=unique(dfb1$Site),  nighthour=hours, behavior = unique(dfb1$behavior)) 
#----expanded data, 257664 obs of 4 vars 

temp$night <-  as.POSIXct(format(temp$nighthour,"%Y-%m-%d"),tz="UTC")

temp2<-merge(temp,newnights) #42600 observations
summary(temp2$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#  3150  3150  2634  1977  2679  2157  2679  2697  2697  2679  2652  2679  2697  2679  2697  2697 

temp3<-merge(temp2,dfb1, all.x=TRUE)
dim(temp3) # 42600 obs of 6 vars 
summary(temp3) #(previously 64416) NA obs - 38525 NAs added 

temp4 <- temp3 %>% mutate_at(6, ~replace_na(.,0))
sum(temp4$batpass) #19438 
summary(temp4)

# Trim to match active.hours 
temp5 <- merge(temp4, nighthours, all.y = TRUE) # 51224 obs 
summary(temp5) # 8624 nighthours added as NAs 


temp5$night <- as.Date(temp5$nighthour)
temp6 <- temp5 %>% select(-active.night) 
temp7 <- temp6 %>% mutate_at(5, ~replace_na(.,0)) # make NA bat passes 0 
summary(temp7) 
sum(temp7$batpass) # 19438
summary(temp7$Site)
# Met45 Met95   N02   N04   N08   N09   N10   N11   N14   P02   P04   P08   P09   P10   P11   P14 
#  3780  3780  3172  2374  3226  2590  3226  3238  3238  3226  3208  3226  3238  3226  3238  3238 

summary(temp7) # 51224 obs of 5 vars
head(temp7)
#    Site           nighthour  behavior      night batpass
# 1 Met45 2020-07-16 08:00:00 Commuting 2020-07-16       0
# 2 Met45 2020-07-16 08:00:00   Feeding 2020-07-16       0
# 3 Met45 2020-07-16 08:00:00    Social 2020-07-16       0
# 4 Met45 2020-07-16 09:00:00 Commuting 2020-07-16       0
# 5 Met45 2020-07-16 09:00:00   Feeding 2020-07-16       0
# 6 Met45 2020-07-16 09:00:00    Social 2020-07-16       0


```

# Only turbines 8 and 11 
Combine with insects

```{r}
#filter bat passes to only include feeding passes from turbines 8 and 11 
temp8 <- temp7 %>% 
  filter(Site %in% c("N08", "P08", "N11", "P11"), behavior == "Feeding") %>% 
  droplevels() 
# 3584 observations 

sum(temp8$batpass)# 426 feeding passes 

bugs <- bugs %>% select(-X) # remove the index column 
names(bugs)
# "Date"          "Time"          "Hour"          "Count"         "InsectPresAbs" "Site"          "Viable"   

str(bugs)
dim(bugs)
# 11420    7

bugs$Site <- as.factor(bugs$Site)
bugs$Viable <- as.factor(bugs$Viable)
bugs$date <- as.Date(bugs$Date, format = "%m.%d.%y")
bugs$datetime <- paste(bugs$date, bugs$Time, sep = " ")
bugs$datetime <- ymd_hms(bugs$datetime)
bugs1 <- bugs %>% mutate(nighttime = datetime - 12*60*60,
                         night = as.Date(nighttime),
                         jnight = yday(night)) %>% select(-Hour) 
bugs2 <- bugs1 %>% mutate(date.hour = as.character(datetime), 
                           night.hour = as.character(nighttime),
                           HOUR.12 = hour(night.hour), 
                           HOUR = hour(date.hour)) %>% select(-c(date.hour, night.hour))
tz(bugs2$datetime) # UTC
tz(bugs2$nighttime) # UTC

bugs2$datetime <- force_tz(bugs2$datetime, tzone = "Europe/Berlin")
tz(bugs2$datetime) # Europe / Berlin

bugs2$nighttime <- force_tz(bugs2$nighttime, tzone = "Europe/Berlin")
tz(bugs2$nighttime) # Europe / Berlin 

summary(bugs2)

bugs2$nighthour <- paste(bugs2$night, " ",bugs2$HOUR.12, ":00", sep= "")
bugs2$nighthour <- ymd_hm(bugs2$nighthour)
bugs2$datehour <- paste(bugs2$date, " ",bugs2$HOUR, ":00", sep= "")
bugs2$datehour <- ymd_hm(bugs2$datehour)

tz(bugs2$datehour) # UTC
tz(bugs2$nighthour) # UTC

bugs2$datehour <- force_tz(bugs2$datehour, tzone = "Europe/Berlin")
tz(bugs2$datehour) # Europe / Berlin

bugs2$nighthour <- force_tz(bugs2$nighthour, tzone = "Europe/Berlin")
tz(bugs2$nighthour) # Europe / Berlin 

summary(bugs2)
# write.csv(bugs2, "insect data hourly with jnight hour12.csv")

# replace NAs with zeros - these are when the camera was active but no data was collected/the image was not viable
bugs3 <- bugs2 %>% select(Site, nighthour, Count) 
levels(bugs3$Site) <- c("N08", "N11",  "P08", "P11") # rename sites
bugs3[is.na(bugs3)] <- 0
summary(bugs3)
sum(bugs3$Count)# 4563 bugs 

## How many images had at least one insect in them? 

# 10642 viable images that contained no insects or that occurred outside of the bat recording period. 
justbugs <- bugs3 %>% filter(Count!=0) #578
nobugs <- bugs3 %>% filter(Count==0) #10842

## aggregate to hour 
bugs4 <- bugs3 %>% group_by(Site, nighthour) %>% dplyr::summarize(bugcount = sum(Count)) 
summary(bugs4)
sum(bugs4$bugcount)# 4563 bugs - good! 

temp9 <- temp8 %>% select(-night)
summary(bugs4)
summary(temp8)

batsbugs <- merge(bugs4, temp9, all.x = TRUE) %>% distinct() # 2045 observations of 5 vars

summary(batsbugs) # 2045 observations , 4 NAs 
test <- drop_na(batsbugs) # 2041 obs 
sum(test$bugcount) # no insect counts lost 

batsbugs1 <- test

sum(batsbugs1$bugcount) # 4563 bugs
sum(batsbugs1$batpass)  # 373 bat feeding passes that match up with active camera trap hours. 
summary(batsbugs1$Site)
# N08 N11 P08 P11 
# 442 544 454 601 
# 2041 obs of 5 vars 
```


## Add back temporal and spatial columns, merge with weather  
```{r}

# Site info 
batsbugs1<- left_join(batsbugs1, bat_sitemap) %>% droplevels()
summary(batsbugs1)

# HOUR.12, datehour, date, HOUR, jnight 
batsbugs2 <- batsbugs1 %>% mutate(datehour = nighthour + 12*60*60,
                          date = as.Date(datehour),
                          night = as.Date(nighthour),
                          jnight = yday(night)) 

batsbugs3 <- batsbugs2 %>% mutate(date.hour = as.character(datehour), 
                           night.hour = as.character(nighthour),
                           HOUR.12 = hour(night.hour), 
                           HOUR = hour(date.hour)) %>% select(-c(date.hour, night.hour))

summary(batsbugs3)# 2045 observations 

# write.csv(batsbugs3, "bats and insects combined no weather.csv")

summary(weather.hours)
weather.hours1 <- weather.hours %>% select(nighthour, wind, temp) # 2194 obs of 3 vars 
tz(weather.hours1$nighthour) # UTC

weather.hours1$nighthour <- force_tz(weather.hours1$nighthour, tzone = "Europe/Berlin")
tz(weather.hours1$nighthour) # "Europe/Berlin" 

batsbugs4 <- left_join(batsbugs3, weather.hours1, by = "nighthour") 
summary(batsbugs4) # 4 NAs introduced

batsbugs5 <- drop_na(batsbugs4)
summary(batsbugs5) # 2041 observations 
sum(batsbugs5$batpass) # 372, lost one bat pass 
sum(batsbugs5$bugcount) # 4546, lost 17 insect counts 

# write.csv(batsbugs5, "bats and insects combined with weather.csv")

```
