
---
title: "4. Insect - bat models and figures"
output: html_document
date: "2023-05-23"
---

# Input: Bats and insect data aggregated to night, with and without weather data attached 
# Output - GAM models predicting bat feeding buzzes in response to insect presence/absence, with and without weather, table for appendix describing each model and a figure expressing bat feeding activity in response to hours after sunset. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Set up work space 
```{r}

library(tidyverse)
library(ggplot2)
library(gratia)
library(DHARMa)
library(mgcv)
library(tidymv)
library(knitr)
library(kableExtra)
library(tidycat)
library(broom)
library(ggpubr)
library(cowplot)
library(suncalc)
library(lubridate)

source(file = "~/1. PhD_Main/GitHub_link/HighstatLibV11.R")  
source(file = "~/1. PhD_Main/GitHub_link/HighstatLibV13.R") 

input <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Input"

dataset1 <- "bats and insects combined no weather.csv"
dataset2 <- "bats and insects combined with weather.csv"
dataset3 <- "Marker_night_inventory_0704.2022_edit.csv" 


path1 <- str_c(input, "/", dataset1)
path2 <- str_c(input, "/", dataset2)
path3 <- str_c(input, "/", dataset3)

batsbugs <- read_csv(path1) # 2041 bat-bug hours with 15 vars
batsbugsweather <- read.csv(path2) # 2037 bat-bug hours with 17 vars
active.nights <- read.csv(path3, sep = ";") # # 951 obs of 3 variables

# # for Reed
output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/2. Marker 2019-2020/Marker 2022/SecondDraftAnalyses/Reed/Outputs"

## 
 file.name <- "4. Insect - bat models and figures"
# 
 todays_date <- Sys.Date()
# 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
 dir.name
# 
 dir.create(dir.name) # be careful not to run over an already created directory 

output_today <- dir.name

setwd(output_today)

getwd()

```


## Prepping dataset without weather 
```{r}
#---QUICK INITIAL LOOK AT THE 'batsbugs' DATA OBJECT
str(batsbugs)
head(batsbugs)
tail(batsbugs)
dim(batsbugs)
summary(batsbugs) 
names(batsbugs)

sum(batsbugs$bugcount) # 4563
sum(batsbugs$batpass) # 373 

tz(batsbugs$nighthour) # UTC
tz(batsbugs$datehour) # UTC

batsbugs$datehour <- force_tz(batsbugs$datehour, tzone = "Europe/Berlin")
tz(batsbugs$datehour)

batsbugs$nighthour <- force_tz(batsbugs$nighthour, tzone = "Europe/Berlin")
tz(batsbugs$nighthour)

####################################################

#remove rows with NAs (if any)

batsbugs2 <- na.omit(batsbugs) 
dim(batsbugs2)
# 2041   15 No NAs

#What is the percentage of zero values in bugcount?
100* sum(batsbugs$bugcount == 0) / nrow(batsbugs) 
#[1] 86.6242

100* sum(batsbugs$batpass == 0) / nrow(batsbugs) 
#[1] 94.90446

#add new column that cuts 'bugcount' into three categories
range(batsbugs2$bugcount)
batsbugs3 <- batsbugs2 %>% mutate(category=cut(bugcount, breaks=c(-Inf, 0, 9, Inf), labels=c("no","low","swarm")))
table(batsbugs3$category)

 #   no   low swarm 
 # 1768   176    97 

plot(table(batsbugs3$category)) #very few observations in both category swarm and category swarm

#add new column that cuts 'bugcount' into binary variable
InsectBat <- batsbugs3 %>% mutate(binary=cut(bugcount, breaks=c(-Inf, 0, Inf), labels=c("No insects","Insects active")))
table(InsectBat$binary)
    # No insects Insects active 
    #       1768            273 

plot(table(InsectBat$binary)) #probably use this as predictor

#rename
InsectBat$InsectCount <- InsectBat$bugcount

InsectBat <- InsectBat %>% 
  mutate(binary=cut(InsectCount, breaks=c(-Inf, 0, Inf), 
                    labels=c("No insects","Insects active")),
         binarybat=cut(batpass, breaks=c(-Inf, 0, Inf), 
                    labels=c("No bats","Bats active"))) 

#make sure other categorical variables are factors
InsectBat$fSite <- factor(InsectBat$Site)
InsectBat$fLocality <- factor(InsectBat$Locality)
InsectBat$fHabitat <- factor(InsectBat$Habitat)
InsectBat$binary <- factor(InsectBat$binary)
InsectBat$binarybat <- factor(InsectBat$binarybat)

head(InsectBat)
tail(InsectBat)
summary(InsectBat)
sum(InsectBat$batpass)

#write.csv(InsectBat, "R export bats insects hr no weather.csv") 

par(mfrow = c(1,2), mar = c(5,5,2,2))
boxplot(batpass ~ category, data = InsectBat) 
# based on IR definitions of swarms 

boxplot(batpass ~ binary, data = InsectBat) 
# We do not have enough photos with high insect counts to get much from those distinctions

par(mfrow = c(2,1), mar = c(4,5,1,2), cex.lab = 1.5)
plot(InsectBat$HOUR.12, InsectBat$batpass, xlab= "HOUR.12", ylab = "Bat pass per hour")
plot(InsectBat$HOUR.12, InsectBat$InsectCount, xlab= "HOUR.12", ylab = "Insect count per hour")

summary_insectbatnight <- InsectBat %>% group_by(Site, Locality, night) %>% dplyr::summarize(n = sum(n())) # number of observations per night per site 

# number of hours recorded per site 
summary_insectbathour <- InsectBat %>% group_by(Site, Locality) %>% dplyr::summarize(n = sum(n())) 
summary_insectbathour

#   Site  Locality      n
#   <chr> <chr>     <int>
# 1 N08   Turbine8    442
# 2 N11   Turbine11   544
# 3 P08   Turbine8    454
# 4 P11   Turbine11   601

sum(InsectBat$InsectCount)

bathours <- InsectBat %>% dplyr::filter(batpass>0)

```


## Prepping dataset with weather 
```{r}
#---QUICK INITIAL LOOK AT THE 'batsbugsweather' DATA OBJECT
str(batsbugsweather)
head(batsbugsweather)
tail(batsbugsweather)
dim(batsbugsweather)
summary(batsbugsweather) 
names(batsbugsweather)

batsbugsweather$nighthour <- as.POSIXct(batsbugsweather$nighthour, format = "%Y-%m-%d %H:%M:%OS")
batsbugsweather$datehour <- as.POSIXct(batsbugsweather$datehour, format = "%Y-%m-%d %H:%M:%OS")

tz(batsbugsweather$nighthour) # ""
tz(batsbugsweather$datehour) # "" 

batsbugsweather$datehour <- force_tz(batsbugsweather$datehour, tzone = "Europe/Berlin")
tz(batsbugsweather$datehour)

batsbugsweather$nighthour <- force_tz(batsbugsweather$nighthour, tzone = "Europe/Berlin")
tz(batsbugsweather$nighthour)

####################################################

#remove rows with NAs (if any)
batsbugsweather2 <- na.omit(batsbugsweather) 
dim(batsbugsweather2)
# 2037      17 No NAs

#What is the percentage of zero values in bugcount?
100* sum(batsbugsweather$bugcount == 0) / nrow(batsbugsweather) 
#[1] 86.69612

100* sum(batsbugsweather$batpass == 0) / nrow(batsbugsweather) 
#[1] 94.94354

sum(batsbugsweather$bugcount)# 4546
sum(batsbugsweather$batpass) # 372

#add new column that cuts 'bugcount' into three categories
range(batsbugsweather2$bugcount) # 0 258
batsbugsweather3 <- batsbugsweather2 %>% mutate(category=cut(bugcount, breaks=c(-Inf, 0, 9, Inf), labels=c("no","low","swarm")))

table(batsbugsweather3$category)
plot(table(batsbugsweather3$category)) #very few observations in both category swarm and category swarm

#add new column that cuts 'bugcount' into binary variable
InsectBatW <- batsbugsweather3 %>% mutate(binary=cut(bugcount, breaks=c(-Inf, 0, Inf), labels=c("No insects","Insects active")))
table(InsectBatW$binary)
plot(table(InsectBatW$binary)) #probably use this as predictor

#rename
InsectBatW$InsectCount <- InsectBatW$bugcount

InsectBatW <- InsectBatW %>% 
  mutate(binary=cut(InsectCount, breaks=c(-Inf, 0, Inf), 
                    labels=c("No insects","Insects active")))

#make sure other categorical variables are factors
InsectBatW$fSite <- factor(InsectBatW$Site)
InsectBatW$fLocality <- factor(InsectBatW$Locality)
InsectBatW$fHabitat <- factor(InsectBatW$Habitat)
InsectBatW$binary <- factor(InsectBatW$binary)
InsectBatW$night <- as.Date(InsectBatW$night)

head(InsectBatW)
tail(InsectBatW)

#write.csv(InsectBatW, "R export bats insects hr with weather.csv") 

par(mfrow = c(1,2), mar = c(5,5,2,2))
boxplot(batpass ~ category, data = InsectBatW) 
# based on IR definitions of swarms 

boxplot(batpass ~ binary, data = InsectBatW) 
# We do not have enough photos with high insect counts to get much from those distinctions

par(mfrow = c(2,1), mar = c(4,5,1,2), cex.lab = 1.5)
plot(InsectBatW$HOUR.12, InsectBatW$batpass, xlab= "HOUR.12", ylab = "Batpass per hour")
plot(InsectBatW$HOUR.12, InsectBatW$InsectCount, xlab= "HOUR.12", ylab = "InsectCount per hour")

summary_insectbatnight1 <- InsectBatW %>% group_by(Site, Locality, night) %>% dplyr::summarize(n = sum(n())) # number of observations per night per site 

# number of hours recorded per site 
summary_insectbathour1 <- InsectBatW %>% group_by(Site, Locality) %>% dplyr::summarize(n = sum(n()))


summary_insectbathour1
#   Site  Locality      n
#   <chr> <chr>     <int>
# 1 N08   Turbine8    440
# 2 N11   Turbine11   544
# 3 P08   Turbine8    452
# 4 P11   Turbine11   601

summary_insectbathour # compared to hours not aggregated to weather
#  Site  Locality      n
#   <chr> <chr>     <int>
# 1 N08   Turbine8    442
# 2 N11   Turbine11   544
# 3 P08   Turbine8    454
# 4 P11   Turbine11   601

summary(InsectBat$night)
#         Min.      1st Qu.       Median         Mean      3rd Qu.         Max. 
# "2020-07-16" "2020-08-13" "2020-08-30" "2020-08-27" "2020-09-14" "2020-09-28"

summary(InsectBatW$night)
#         Min.      1st Qu.       Median         Mean      3rd Qu.         Max. 
# "2020-07-16" "2020-08-13" "2020-08-30" "2020-08-27" "2020-09-14" "2020-09-28" 

summary(summary_insectbathour$n)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 442.0   451.0   500.0   511.2   560.2   603.0 

```

## Create the hour after sunset column 
```{r}
head(InsectBatW)
head(InsectBat)
head(active.nights)

# manually add 09.30 
thenights <- as.data.frame(active.nights$night %>% unique())
colnames(thenights)[1] = "date"
thenights$date <- as.Date(thenights$date, format = "%d.%m.%Y")
day0930 <- as.data.frame(as.Date("2020-09-30")) 
colnames(day0930 )[1] = "date"
thenights1 <- full_join(thenights, day0930)
# Will have to use night as a proxy for date
# The code below works best if the data variable is named "date" 


#Use location of T-08
photoperiod <- getSunlightTimes(
  date = thenights1$date, 
  keep = c("sunrise", "sunset"),
  lat = 59.49964,
  lon = 11.73199, 
  tz = "Europe/Berlin")

photoperiod1 <- photoperiod %>%
  mutate(
    date = as.POSIXct(date),
    day_length = as.numeric(sunset - sunrise)) %>% distinct()

photoperiod2 <- photoperiod1 %>% mutate(night = as.Date(date)) %>% select(-date) %>% distinct()
summary(photoperiod2)

head(photoperiod2)

head(InsectBat)


tz(InsectBat$datehour) # "Europe/Berlin" 
tz(InsectBatW$datehour) # "Europe/Berlin" 
tz(photoperiod$sunset) # "Europe/Berlin" 

#Combine photoperiod data with InsectBat data
InsectBat1 <- left_join(InsectBat, photoperiod2)
head(InsectBat1)

tz(InsectBat1$nighthour) # "Europe/Berlin"
tz(InsectBat1$sunset) # "Europe/Berlin" 


# Find difference between datehour and sunset 
InsectBat1$difhour <- difftime(InsectBat1$datehour, InsectBat1$sunset, tz == 'Europe/Berlin', unit="hours")

test <- InsectBat1 %>% select(datehour, sunset, difhour)
# Looks good! 

# Hour difference to sunset - variable for modelling 
InsectBat1$HOUR.SS <-  time_length(InsectBat1$difhour,unit="hours")

```


## More data exploration 
```{r}
#Make a new column in the dataset
InsectBat1$batpass30  <- InsectBat1$batpass
#Make a new column in the dataset
InsectBat1$InsectCount30 <-InsectBat1$InsectCount

InsectBat1$InsectCount5 <-InsectBat1$InsectCount

# reassign batpass_sum values over 10 to 10
InsectBat1$batpass30[InsectBat1$batpass>30] <- 30
# reassign InsectCount values over 30 to 30
InsectBat1$InsectCount30[InsectBat1$InsectCount>30] <- 30

# reassign InsectCount values over 10 to 10
InsectBat1$InsectCount5[InsectBat1$InsectCount>5] <- 5

summary(InsectBat1$InsectCount30)

par(mfrow = c(2,1), mar = c(4,5,1,2), cex.lab = 1.5)
plot(InsectBat1$HOUR.12, InsectBat1$batpass30, xlab= "HOUR.12", ylab = "batpass.max.10")
plot(InsectBat1$HOUR.12, InsectBat1$InsectCount30, xlab= "HOUR.12", ylab = "insectcount.max.50")


#What is the percentage of values > 10 in batpass?
100* sum(InsectBat1$bugcount > 10) / nrow(InsectBat1) 
#[1] 4.507594


#What is the percentage of values > 30 in batpass?
100* sum(InsectBat1$InsectCount > 30) / nrow(InsectBat1) 
#[1] 1.714846

#Sketching relationships 


## Relationship between feeding buzzes and insect counts
InsectBat1$logSumIns  <- log1p(InsectBat1$InsectCount) #log x+1 transform insect variable

par(mfrow = c(2,1), mar = c(4,5,1,2), cex.lab = 1.5)
plot(InsectBat1$logSumIns,InsectBat1$batpass)
plot(InsectBat1$InsectCount,InsectBat1$batpass)
#Hmmm - if anything - it seems like there is a negative relationship - use the binary variable as predictor instead!


```


## fit GAM - HOUR.SS (hours difference to sunset) response 
```{r}
summary(InsectBat1)
M1 <- gam(batpass ~ binary + s(HOUR.SS, bs = "gp", k = 15) + s(fLocality, bs="re"),              
          data = InsectBat1, method = "REML",    #REML is not default, but is highly recommended by experts
          family = nb(), select=TRUE)             
summary(M1)

# Family: Negative Binomial(0.059) 
# Link function: log 
# 
# Formula:
# batpass ~ binary + s(HOUR.SS, bs = "gp", k = 15) + s(fLocality, 
#     bs = "re")
# 
# Parametric coefficients:
#                      Estimate Std. Error z value Pr(>|z|)    
# (Intercept)           -4.4976     0.7460  -6.029 1.65e-09 ***
# binaryInsects active   1.6080     0.3931   4.090 4.31e-05 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# Approximate significance of smooth terms:
#                 edf Ref.df Chi.sq p-value    
# s(HOUR.SS)   7.8206     14 65.495 < 2e-16 ***
# s(fLocality) 0.9016      1  8.674 0.00192 ** 
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# R-sq.(adj) =  0.0202   Deviance explained = 40.7%
# -REML = 547.57  Scale est. = 1         n = 2041


par(mfrow = c(2,2))
gam.check(M1, rep=500) #k-index a bit too low, but we'll accept this
gam.check(M1)

overdispersion.M1 <- sum( residuals(M1, "pearson")^2 ) / M1$df.residual
overdispersion.M1
#[1] 0.9576367

dotchart(InsectBat$InsectCount)
hist(InsectBat$InsectCount)
summary(InsectBat$InsectCount)
sd(InsectBat$InsectCount)

pdata <- with(InsectBat1,
              expand.grid(binary = levels(binary), 
                          InsectCount = seq(min(InsectCount), max(InsectCount), length=100),
                          fLocality = c("Turbine11"),
                          HOUR.SS = seq(min(HOUR.SS), max(HOUR.SS), length=100))) 
#Note - it does not matter whether you choose T11 or T08 - the plots look the same
head(pdata)
tail(pdata)

fit <- data.frame(predict(M1, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (1*se.fit), lower = fit-(1*se.fit)) #1*se.fit instead of two because huge variation
pred <- cbind(pdata,fit)

head(pred)
summary(pred)


```


## # "Bats - Insects" model 
fit GAM - HOUR.SS (since sunset) response 
```{r}
summary(InsectBat1)
M1 <- gam(batpass ~ binary + s(HOUR.SS, bs = "gp", k = 15) + s(fLocality, bs="re"),              
          data = InsectBat2, method = "REML",    #REML is not default, but is highly recommended by experts
          family = nb(), select=TRUE)             
summary(M1)

# Family: Negative Binomial(0.063) 
# Link function: log 
# 
# Formula:
# batpass ~ binary + s(HOUR.SS, bs = "gp", k = 15) + s(fLocality, 
#     bs = "re")
# 
# Parametric coefficients:
#                      Estimate Std. Error z value Pr(>|z|)    
# (Intercept)           -4.6427     0.8040  -5.774 7.72e-09 ***
# binaryInsects active   1.5868     0.3822   4.152 3.30e-05 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# Approximate significance of smooth terms:
#                 edf Ref.df Chi.sq p-value    
# s(HOUR.SS)   8.7132     14 80.267 < 2e-16 ***
# s(fLocality) 0.9061      1  9.099 0.00153 ** 
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# R-sq.(adj) =  0.0141   Deviance explained = 43.9%
# -REML = 544.27  Scale est. = 1         n = 2045

par(mfrow = c(2,2))
gam.check(M1, rep=500) #k-index a bit too low, but we'll accept this
gam.check(M1)

overdispersion.M1 <- sum( residuals(M1, "pearson")^2 ) / M1$df.residual
overdispersion.M1
#[1] 0.9378456


#M1 <- gam(batpass ~ binary + s(HOUR.12, by = fLocality, bs = "gp", k = 15) ,              
#          data = InsectBat, method = "REML",                  #REML is not default, but is highly recommended by experts
#          family = nb(), select=TRUE)             
#summary(M1)

#par(mfrow = c(2,2))
#gam.check(M1, rep=500) 

#overdispersion.M1 <- sum( residuals(M1, "pearson")^2 ) / M1$df.residual
#overdispersion.M1
#[1] 0.7205232

#M1 also look OK, but underdispersed
#Use M1, also because that is how we dealt with fLocility in the other GAMs (per night)

##Sketching the results - M1
dotchart(InsectBat2$InsectCount)
hist(InsectBat2$InsectCount)
summary(InsectBat2$InsectCount)
sd(InsectBat2$InsectCount)

pdata <- with(InsectBat2,
              expand.grid(binary = levels(binary), 
                          InsectCount = seq(min(InsectCount), max(InsectCount), length=100),
                          fLocality = c("Turbine11"),
                          HOUR.SS = seq(min(HOUR.SS), max(HOUR.SS), length=100))) 
#Note - it does not matter whether you choose T11 or T08 - the plots look the same
head(pdata)
tail(pdata)

fit <- data.frame(predict(M1, newdata=pdata, se.fit=TRUE, type = 'response'))
fit <- transform(fit, upper = fit + (1*se.fit), lower = fit-(1*se.fit)) #1*se.fit instead of two because huge variation
pred <- cbind(pdata,fit)

head(pred)
summary(pred)

```


# Exploratory figures 
```{r}

## Survey effort
InsectBat1$active.hour <- "TRUE"

library(MetBrewer)
names(InsectBat1)
levels(InsectBat1$binary)
levels(InsectBat1$binarybat)
# Appendix 

summary(InsectBat1$night)

bugsonly <- InsectBat1 %>% filter(binary == "Insects active") %>% droplevels()
# 273 trap hours
summary(bugsonly$night)
#         Min.      1st Qu.       Median         Mean      3rd Qu.         Max. 
# "2020-07-16" "2020-08-08" "2020-08-16" "2020-08-18" "2020-08-27" "2020-09-28" 
batsonly <- InsectBat1 %>% filter(binarybat == "Bats active") %>% droplevels()
# 104 trap hours 
summary(batsonly$night)
#         Min.      1st Qu.       Median         Mean      3rd Qu.         Max. 
# "2020-07-18" "2020-08-03" "2020-08-12" "2020-08-12" "2020-08-18" "2020-09-18" 

ggplot() + 
  geom_point(data = bugsonly, aes(x = night, y = binary), 
             color = "#E67423",  size = 4, alpha = 0.5)  +
  geom_point(data = batsonly, aes(x = night, y = binarybat), 
             color = "#999999", alpha = 0.5, size = 4) +
  geom_point(data = InsectBat1, aes(x = night, y = active.hour), 
             alpha = 0.5, size = 4, color = "#23293F") + 
  facet_wrap(~Site) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), 
text = element_text(size = 20),
axis.line = element_line(colour = "black"), 
legend.position="none") + 
  xlab("") + ylab("") + 
  scale_x_date(limits = c(min = as.Date("2020-07-01"), 
                          max = as.Date("2020-10-01"), limits = c(0,0))) + 
    scale_y_discrete(labels = c("Bats active", 
    "Insects active", "Camera - detector active"))
  
```


### Figures for the paper 
```{r}

reference = c("#B40F20" , "#FCC200", "#46ACC8", "#999999") 


bugsonly11 <- bugsonly %>% filter(Locality == "Turbine11") # 143 obs 

plt2 <- 
  ggplot(pred, aes(x=HOUR.SS, y = fit, group = factor(binary))) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'lightgrey', alpha = 0.5) +
  geom_line(aes(color = binary), linewidth = 2) + 
  scale_color_manual(
    values = c("No insects" = "black" , "Insects active" = "#E67423")) +
  labs(x = "Time", y ="Predicted bat feeding passes (lines) / Insect counts (dots)") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 20),
        legend.position="bottom", 
        axis.line = element_line(colour = "black")) +
  theme(text = element_text(size = 15)) +
  geom_count(data = bugsonly11, 
             aes(x = HOUR.SS, y = InsectCount5), color = "#E67423", alpha = 0.5) +
  scale_size_continuous(name=" ") + 
 labs(color  = " ") 
plt2

# Figure 
plt2 <- 
  ggplot(pred, aes(x=HOUR.SS, y = fit, group = factor(binary))) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = 'lightgrey', alpha = 0.5) +
  geom_line(aes(color = binary), linewidth = 2) + 
  scale_color_manual(
    values = c("No insects" = "black" , "Insects active" = "#E67423")) +
  labs(x = "Hours since sunset", y ="Bat feeding passes per hour") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 20),
        legend.position="bottom", 
        axis.line = element_line(colour = "black")) +
  theme(text = element_text(size = 15)) +
  scale_size_continuous(name=" ") + 
 labs(color  = " ") 
plt2

```


# "Insects" Model
## Evaluate influence of weather on insect data ("Insects" model)

```{r}
head(InsectBatW) 

M2 <- gam(InsectCount ~ s(jnight, bs = "gp", k = 60) + s(wind) + s(temp) + s(fLocality, bs="re"),              
          data = InsectBatW, method = "REML",    #REML is not default, but is highly recommended by experts
          family = nb(), select=TRUE)             
summary(M2)

# Family: Negative Binomial(0.048) 
# Link function: log 
# 
# Formula:
# InsectCount ~ s(jnight, bs = "gp", k = 60) + s(wind) + 
#     s(temp) + s(fLocality, bs = "re")
# 
# Parametric coefficients:
#             Estimate Std. Error z value Pr(>|z|)
# (Intercept)  -0.3718     0.6676  -0.557    0.578
# 
# Approximate significance of smooth terms:
#                  edf Ref.df Chi.sq  p-value    
# s(jnight)    10.8231     59 89.026   0.0102 *  
# s(wind)       1.7224      9  4.908   0.0371 *  
# s(temp)       5.2479      9 62.505  < 2e-16 ***
# s(fLocality)  0.9555      1 21.072 1.06e-06 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# R-sq.(adj) =  0.038   Deviance explained = 24.9%
# -REML = 1682.6  Scale est. = 1         n = 2037

gam.check(M2)
par(mfrow = c(2,2))
gam.check(M2, rep=500) #k-index a bit too low, but we'll accept this
gam.check(M2)

overdispersion.M2 <- sum( residuals(M2, "pearson")^2 ) / M2$df.residual
overdispersion.M2
#[1] 1.02571

```



# M1 and M2 results tables
```{r}

## "Bats - Insects" model 
options(knitr.table.format = "html") 
 # c("Locality", "Habitat: Natural x Julian night", " Habitat: Turbine pad x Julian night", "Average nightly wind speed")


### M1 - Insect2
# Table of the parametric effects
 
mpara <- tidy(M1, parametric = TRUE, conf.int = TRUE) %>% 
  mutate_if(is.numeric, format, digits=3) 

colnames(mpara) <- c("Parametric terms", "Estimate", "SE", "z-value", "P", "CI (LL)", "CI (UL)")

mpara$"Parametric terms" <- gsub("binaryinsects active",
                                 "Insects active (binary)",
                                 mpara$"Parametric terms")

summary(M1)
mparatab <- kable(mpara)  %>%
  kable_minimal() %>% 
  row_spec(0, italic = TRUE) 
mparatab


# Table of the smooth effects
msmooth <- tidy(M1, parametric = FALSE) %>% 
  mutate_if(is.numeric, format, digits=3) 

colnames(msmooth) <- c("Smooth terms", "Estimate", "edf", "Chi.sq", "P") 
msmooth$`Smooth terms` <- as.factor(msmooth$`Smooth terms`)
levels(msmooth$`Smooth terms`) <- (c("Hour","Locality" ))

summary(M1)
msmoothtab <- kable(msmooth) %>% 
  kable_minimal() %>% 
  row_spec(0, italic = TRUE) 
msmoothtab

### M2 
# Table of the parametric effects
 
mpara <- tidy(M2, parametric = TRUE, conf.int = TRUE) %>% 
  mutate_if(is.numeric, format, digits=3) 

colnames(mpara) <- c("Parametric terms", "Estimate", "SE", "z-value", "P", "CI (LL)", "CI (UL)") 

mparatab <- kable(mpara)  %>%
  kable_minimal() %>% 
  row_spec(0, italic = TRUE) 
mparatab


# Table of the smooth effects
msmooth <- tidy(M2, parametric = FALSE) %>% 
  mutate_if(is.numeric, format, digits=3) 

colnames(msmooth) <- c("Smooth terms", "Estimate", "edf", "Chi.sq", "P") 
msmooth$`Smooth terms` <- as.factor(msmooth$`Smooth terms`)
levels(msmooth$`Smooth terms`) <- (c(
  "Julian night",
  "Wind speed", "Temperature", "Locality" ))


msmoothtab <- kable(msmooth) %>% 
  kable_minimal() %>% 
  row_spec(0, italic = TRUE) 
msmoothtab

```

